Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic

### 2022-04-06 psy_peds_3_LURIE.Rmd

---
title: "Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 72
---

@@

listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "lubridate")
paket(listOfPackages)
```

@@

@@

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "/Users/hanauer/Desktop/DONT_INDEX/4ceData/Input/phase22_11NOV2021/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
data_update_date <- "2021-11-04"
time_period <-"months" #possible values: days, weeks, months
```

```{r}
# cut-off dates used in the analysis, these should not be changed
# history_start_date <- as.Date("2019-01-01")
# history_end_date <- as.Date("2021-05-31")
pandemic_start_date <- as.Date("2020-03-15")
start_date_plots <- as.Date("2018-12-01")
end_date_plots <- as.Date("2021-07-01")
```

@@

```{r message=FALSE, warning=FALSE}
### Read the CSV input files
source("R/readInputFiles.R")

@@

### Read the ICD10 codes file

@@

@@

@@

```{r}
# Filtering on inpatients here (in_hospital == 1), to reduce subsequent compute time of the encounter length calculation 
clinical_raw <- filter(clinical_raw, in_hospital == 1)
```

@@

```{r}
count_sequences_hospitalisation <- function(df, ...) {
  seq_hospitalisation_df <- data.frame(total_span = seq(min(df$days_since_admission),
                                                        max(df$days_since_admission))
  ) %>%
    left_join(df, by = c("total_span" = "days_since_admission")) %>%
    replace_na(list(in_hospital = 0))
  count_sequences <- rle(seq_hospitalisation_df$in_hospital)
  count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
  n_sequences <- seq_along(count_sequences_1$lengths)
  sequences <- rep.int(n_sequences, count_sequences_1$lengths)
  sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
  stopifnot(length(df$days_since_admission) == length(sequences))
  data.frame(days_since_admission = df$days_since_admission,
             n_hospitalisation = sequences,
             len_hospitalisation = sequences_len)
}
stopifnot(all(clinical_raw$in_hospital == 1))
hospitalisations_seq_df <- clinical_raw %>%
  distinct(patient_num, cohort, days_since_admission, in_hospital) %>%
  group_by(patient_num, cohort) %>%
  group_modify(count_sequences_hospitalisation)

@@

@@

```{r}
### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
codesToReview <- obs_raw %>% dplyr::filter(concept_type == "DIAG-ICD10",
                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", concept_code))
print(codesToReview)

### Calculate how many of the mental codes are in the dataset
mentalHealthCodes <- unique( obs_raw %>%
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 concept_code %in% icdCodes$ICD10_Code ) %>%
  dplyr::mutate( concept_code = as.character( concept_code ), 
                 concept_length = nchar( concept_code ), 
                 concept_check = ifelse( concept_length == 4 | concept_length > 8, "Check", "OK")) %>%
  dplyr::select( concept_code, concept_length, concept_check) )

@@

print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
```

### Calculating age at hospitalization time
```{r}
# Calculating patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age")], by = c("patient_num")) %>% 
  left_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
            by = c("patient_num", "days_since_admission"))
stopifnot(nrow(obs_raw_age) == nrow(obs_raw))
obs_raw_age <- obs_raw_age[ !is.na( obs_raw_age$calendar_date), ]
obs_raw_age$age_time_diagnosis <- obs_raw_age$age - floor(as.numeric(as.Date(data_update_date) - as.Date(obs_raw_age$calendar_date, dateFormat)) / 365.25)
```

@@

``` {r}
obs_filtered_age <- obs_raw_age %>% 
  filter(concept_type == 'DIAG-ICD10', 
         days_since_admission >= 0)

@@

```{r}
# Steps necessary to filter rows:
  # - encounters without ICD
  # - ICD diagnosed outside of encounters
  # - patients from the patients table that don't have any encounters matching the criteria
clinical_table <- clinical_filtered %>% 
  inner_join( distinct(obs_filtered_age, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
  distinct(patient_num, n_hospitalisation) %>% 
  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
# Adding new columns
observation_table <- obs_filtered_age %>% 
  inner_join(clinical_table[, c("patient_num", "days_since_admission")], by = c("patient_num", "days_since_admission"))
demographic_table <- demo_raw %>% 
  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
all_codes <- observation_table %>% 
   inner_join(select(clinical_table, -calendar_date), by = c("patient_num", "days_since_admission")) %>% 
   inner_join(demographic_table[, c("patient_num", "sex")], by = "patient_num")
```

```{r}
stopifnot(unique(all_codes$cohort) == "AllAdm")
stopifnot(unique(all_codes$in_hospital) == 1)
stopifnot(unique(all_codes$len_hospitalisation >= min_hosp_days))
stopifnot(nrow(all_codes) == nrow(observation_table))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(demographic_table$patient_num)))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(observation_table$patient_num)))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(clinical_table$patient_num)))
stopifnot(nrow(distinct(all_codes, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table, patient_num, n_hospitalisation)))
stopifnot(nrow(distinct(all_codes, patient_num, days_since_admission)) == nrow(distinct(observation_table, patient_num, days_since_admission)))
```

@@

```{r}
# Filtering hospitalizations where patient age is in inclusion criteria
observation_table_ado <- filter(observation_table, age_time_diagnosis >= 11 & age_time_diagnosis <= 17)

@@

```{r}
stopifnot(unique(all_codes_ado$cohort) == "AllAdm")
stopifnot(unique(all_codes_ado$in_hospital) == 1)
stopifnot(unique(all_codes_ado$len_hospitalisation >= min_hosp_days))
stopifnot(nrow(all_codes_ado) == nrow(observation_table_ado))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(demographic_table_ado$patient_num)))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(observation_table_ado$patient_num)))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(clinical_table_ado$patient_num)))
stopifnot(nrow(distinct(all_codes_ado, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table_ado, patient_num, n_hospitalisation)))
stopifnot(nrow(distinct(all_codes_ado, patient_num, days_since_admission)) == nrow(distinct(observation_table_ado, patient_num, days_since_admission)))
```

>>>>

---
title: "Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic"
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 72
---

@@

#listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr")
#paket(listOfPackages)

# trying to remove kableExtra to run w/o docker
listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "tsModel", "gridExtra", "dplyr")
paket(listOfPackages)
```

@@

### Set working directory

```{r}
#setwd("~/projects/nu4ce/4ceData_files/Phase2.2.psy_peds_trends")
```

@@

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "../4ce_ped_psych/data"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%m-%y" #"%d-%b-%y"
min_hosp_days <- 0
data_update_date <- "2021-06-05"
time_period <-"months" #possible values: days, weeks, months
```

```{r}
# cut-off dates used in the analysis, these should not be changed
#start_date_plots <- as.Date("2019-01-01")
#history_end_date <- as.Date("2021-05-31")
pandemic_start_date <- as.Date("2020-03-15")
start_date_plots <- as.Date("2018-12-01")
end_date_plots <- as.Date("2021-07-01")
```

@@

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
### Read the CSV input files
source("R/readInputFiles.R")

@@

### Read in Lurie files

```{r}
non_psych_inpt_counts <- read.csv(paste0(folder_4ce_files, "/non_psych_admit_inpt_11_17.csv"))
admit_all_codes <- read.csv(paste0(folder_4ce_files, "/psych_admit_all_codes.csv"))
admit_demog <- read.csv(paste0(folder_4ce_files, "/psych_admit_demog.csv"))
admit_inpt_primary <- read.csv(paste0(folder_4ce_files, "/psych_admit_inpt_primary.csv"))
```

Raw Data Counts

```{r}
nrow(unique(data.frame(admit_all_codes$patient_id)))
nrow(unique(data.frame(admit_demog$patient_id)))
```
### Read the ICD10 codes file

@@

Old codes that Nelson used. These were updated to the ICD code list above in order to remove codes for accidental poisoning

```{r}
icdCodes_old <- read.csv("public-data/psychiatric_ICD-10_codes_adjusted.csv", header = TRUE, colClasses = "character")

colnames(icdCodes_old) <- c("disorder_group", "ICD10_Code", "description")

icdCodes_old <- icdCodes_old %>% 
  mutate(ICD10_Code = gsub("ICD10:", "", ICD10_Code))

excluded_codes <- icdCodes_old %>% anti_join(icdCodes)
qc_check <- admit_all_codes %>% filter(icd_10_code %in% excluded_codes$ICD10_Code) 
nrow(unique(data.frame(qc_check$patient_id)))
```

@@

**Meghan Note: (Skip-to Lurie Data Formatting section)**

@@

```{r eval=FALSE, include=FALSE}
# Filtering on inpatients here (in_hospital == 1), to reduce subsequent compute time of the encounter length calculation 
clinical_raw <- filter(clinical_raw, in_hospital == 1)
```

@@

```{r eval=FALSE, include=FALSE}
count_sequences_hospitalisation <- function(df, ...) {
  seq_hospitalisation_df <- data.frame(total_span = seq(min(df$days_since_admission),
                                                        max(df$days_since_admission))
  ) %>%
    left_join(df, by = c("total_span" = "days_since_admission")) %>%
    replace_na(list(in_hospital = 0))
  count_sequences <- rle(seq_hospitalisation_df$in_hospital)
  count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
  n_sequences <- seq_along(count_sequences_1$lengths)
  sequences <- rep.int(n_sequences, count_sequences_1$lengths)
  sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
  stopifnot(length(df$days_since_admission) == length(sequences))
  data.frame(days_since_admission = df$days_since_admission,
             n_hospitalisation = sequences,
             len_hospitalisation = sequences_len)
}
stopifnot(all(clinical_raw$in_hospital == 1))
hospitalisations_seq_df <- clinical_raw %>%
  distinct(patient_num, cohort, days_since_admission, in_hospital) %>%
  group_by(patient_num, cohort) %>%
  group_modify(count_sequences_hospitalisation)

@@

# Lurie Data Formatting

**Clean up Lurie encounters**

As noted by Nelson: "some patients will have two encounters back-to-back. These should be considered the SAME encounter. The reason for having two encounters recorded is because medical admissions and psych admissions are considered separate admissions in our hospital. A simple way to deal with this redundancy is to check encounters for the same patient that start on the same day as the end of the prior encounter and merge those two into one encounter."

**First we will combine all of the psych + non-psych data pulls**

```{r}
# admit_demog$pysch = 1
# non_psych_inpt_counts$pysch = 0

admit_demog <- rbind(admit_demog, non_psych_inpt_counts)

## appears to be no overlap?
#View(admit_demog %>% distinct(patient_id, encounter_id, pysch) %>% group_by(psych) %>%  mutate(count = n()))
```

```{r}
## format dates
admit_demog$i_hsp_adm_dttm <- strptime(admit_demog$hsp_adm_dttm, format="%m/%d/%Y %H:%M") %>% as.POSIXct()
admit_demog$i_hsp_dsch_dttm <- strptime(admit_demog$hsp_dsch_dttm, format="%m/%d/%Y %H:%M") %>% as.POSIXct()

# # filter ages
# admit_demog <- admit_demog %>% filter(age_years > 10 & age_years < 18)  %>%
#  filter(i_hsp_adm_dttm >= start_date_plots & i_hsp_adm_dttm <= end_date_plots)

# unique encounter numbers should equal the row number (TRUE)
nrow(unique(data.frame(admit_demog$encounter_id))) == nrow(admit_demog)
nrow(admit_demog %>% distinct(patient_id, encounter_id))

encounters <- admit_demog %>% 
  mutate(gender = if_else(gender == "Male", "male", "female")) %>% 
  group_by(patient_id) %>% 
  arrange(patient_id, i_hsp_adm_dttm) %>% 
  filter(i_hsp_adm_dttm >= start_date_plots & i_hsp_adm_dttm <= end_date_plots) %>% 
  mutate(encounter_group = seq_along(encounter_id),
         day_lag = as.numeric(round(difftime(i_hsp_adm_dttm, lag(i_hsp_dsch_dttm), unit = "hours"), 2))) %>% 
  data.frame() %>%
  # adding 100 to temporarily handle the NAs
  mutate(day_lag = ifelse(is.na(day_lag), 100, day_lag),
  # here we want to ensure that if subsequent encounters occurred within a day (24 hours), that we impute the initial encounter_group as per Nelson's suggestion
         i_encounter_group = ifelse(day_lag <= 24, lag(encounter_group), encounter_group),
  consolidated = ifelse(day_lag <= 24, 1, 0))

# this is suggesting that 408 need to get consolidated?
table(encounters$consolidated)

check_lag <- encounters %>% filter(!day_lag == 100)
summary(check_lag$day_lag)
less24 <- check_lag %>% filter(day_lag <= 24) #449
summary(less24$day_lag)
hist(less24$day_lag)

encounters <- encounters %>%
  # in some cases there are more than two back-to-back encounters
  # to filter for these, first we will group by the patient_id and the new encounter group
  group_by(patient_id, i_encounter_group) %>%
  # next, we will calculate the latest admission date of this encounter
  mutate(max_date = max(i_hsp_adm_dttm)) %>%
  ungroup() %>%
  # next, we group just by Patient, and we identify the latest max_date
  # we then calculate the current time from the lag(max_date) and we will identify those within 24 hours as another encounter that should be grouped
  group_by(patient_id) %>%
  mutate(lag_date = lag(max_date),
         time_diff = as.numeric(round(difftime(i_hsp_adm_dttm, lag_date, unit = "hours"), 2))) %>%
  # adding 100 to temporarily handle the NAs
  mutate(time_diff = ifelse(is.na(time_diff), 100, time_diff),
    i_encounter_group = ifelse(time_diff <= 24, lag(i_encounter_group), i_encounter_group),
    consolidated2 = ifelse(time_diff <= 24, 1, 0),
  # recalculate the encounter sequence
  encounter_num = dense_rank(i_encounter_group)) %>% 
  group_by(patient_id, encounter_num)

# this is suggesting that 408 need to get consolidated?
table(encounters$consolidated2)

check_lag <- encounters %>% filter(!time_diff == 100)
summary(check_lag$time_diff)
less24 <- check_lag %>% filter(time_diff <= 24) #397 patients/admissions said to need to be consolidated
summary(less24$time_diff)
hist(less24$time_diff)

encounters <- encounters %>% 
  slice(1L)

nrow(encounters %>% distinct(patient_id, encounter_num))
```

Check to see if we should have combined any additional encounters

```{r}
test = encounters %>% 
       group_by(patient_id, encounter_num) %>% 
       mutate(max_date = max(i_hsp_adm_dttm)) %>% 
  ungroup() %>%
  group_by(patient_id) %>% 
  mutate(lag_date = lag(max_date),
         time_diff = as.numeric(difftime(i_hsp_adm_dttm, lag_date, unit = "hours")),
         check = if_else(time_diff <= 24 & encounter_num != lag(encounter_num), 1, 0))

#View(test %>% filter(check == 1))
```

**Add number of hospitalizations and LOS of stay for each patient**

```{r}
clinical_raw <- encounters %>% 
  group_by(patient_id) %>% 
  mutate(n_hospitalisation = n_distinct(as.character(encounter_num))) %>% 
  group_by(patient_id, encounter_num) %>% 
  # calculate the hospitalization length, while taking care to select the max(i_hsp_dsch_dttm) in the case that we have more than 1 row per the same encounter
  mutate(len_hospitalisation = round(as.numeric(difftime(max(i_hsp_dsch_dttm), min(i_hsp_adm_dttm), unit = "days"))),
         # replace last discharge with the true discharge datetime
         i_hsp_dsch_dttm = max(i_hsp_dsch_dttm),
         # replace age_years with earliest age on admission 
         age_years = min(age_years)) %>% 
  rename("date" = i_hsp_adm_dttm,
         "patient_num" = patient_id) %>% 
  group_by(patient_num, encounter_num) %>% 
  slice(1L) %>% 
  ungroup() %>% 
  select(patient_num, date, i_hsp_dsch_dttm, race, ethnicity, gender, age_years, encounter_num, n_hospitalisation, len_hospitalisation) %>% 
  distinct()

nrow(clinical_raw %>% distinct(patient_num, encounter_num))
```

@@

```{r echo=TRUE}
### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
codesToReview <- admit_all_codes %>% dplyr::filter(icd_10_code == "DIAG-ICD10",
                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", icd_10_code))
print(codesToReview)

### Calculate how many of the mental codes are in the dataset
mentalHealthCodes <- unique( admit_all_codes %>%
  dplyr::filter( #concept_type == "DIAG-ICD10", 
                 icd_10_code %in% icdCodes$ICD10_Code ) %>%
  dplyr::mutate( icd_10_code = as.character( icd_10_code ), 
                 concept_length = nchar( icd_10_code ), 
                 concept_check = ifelse( icd_10_code == 4 | concept_length > 8, "Check", "OK")) %>%
  dplyr::select( icd_10_code, concept_length, concept_check) )

@@

print( paste0("There are ", length(mentalHealthCodes$icd_10_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
```

### Calculating age at hospitalization time

```{r eval=FALSE, include=FALSE}
# Calculating patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age")], by = c("patient_num")) %>% 
  left_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
            by = c("patient_num", "days_since_admission"))
stopifnot(nrow(obs_raw_age) == nrow(obs_raw))
obs_raw_age$age_time_diagnosis <- obs_raw_age$age - floor(as.numeric(as.Date(data_update_date) - as.Date(obs_raw_age$calendar_date, dateFormat)) / 365.25)
```

@@

```{undefined eval=FALSE, include=FALSE}
obs_filtered_age <- obs_raw_age %>% 
  filter(concept_type == 'DIAG-ICD10', 
         days_since_admission >= 0)

@@

### Further filtering Lurie data

```{r}
clinical_raw$date <- strptime(clinical_raw$date, format="%Y-%m-%d") %>% as.POSIXct()

# create list of calendar_dates
date_ids <- seq(as.Date("2019-01-01"), as.Date(end_date_plots), 
                by= "day") %>% 
  data.frame() %>% 
  rename('calendar_date' = '.')

date_ids <- date_ids %>% 
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = as.Date(cut( date, breaks = "week")),
                 month = as.Date(cut( date, breaks = "month")),
                 year = format( date, "%Y"), 
                 period = ifelse( date < pandemic_start_date,
                                  "before_pandemic","during_pandemic"))

clinical_filtered <- clinical_raw %>% 
  filter(len_hospitalisation >= min_hosp_days) %>% 
  left_join(., date_ids, by = "date") %>% 
   dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                  weeks = as.Date(cut( date, breaks = "week")),
                  month = as.Date(cut( date, breaks = "month")),
                  year = format( date, "%Y"), 
                  period = ifelse( date < pandemic_start_date,
                                   "before_pandemic","during_pandemic"))

# check breaks example from BCH
example_bch <- read.delim2("example.txt", header = TRUE, sep = "")
example_lurie <- date_ids %>% distinct(calendar_date, weeks, month, year, period)

example_bch$calendar_date <- as.Date(example_bch$calendar_date)
example_bch$weeks <- as.Date(example_bch$weeks)
example_bch$month <- as.Date(example_bch$month)
example_bch$year <- as.character(example_bch$year)
example_bch <- example_bch %>% filter(!calendar_date >= "2021-06-01")

diff <- example_bch %>% anti_join(example_lurie)
nrow(diff)
diff <- example_lurie %>% anti_join(example_bch)
nrow(diff)
identical(example_bch, example_lurie)
all.equal(example_bch, example_lurie)

if( time_period == "weeks"){
  clinical_filtered$time_p <- clinical_filtered$weeks
}else if(time_period == "days"){
  clinical_filtered$time_p <- clinical_filtered$date
}else if(time_period == "months"){
  clinical_filtered$time_p <- clinical_filtered$month
}
```

```{r eval=FALSE, include=FALSE}
# Steps necessary to filter rows:
  # - encounters without ICD
  # - ICD diagnosed outside of encounters
  # - patients from the patients table that don't have any encounters matching the criteria
clinical_table <- clinical_filtered %>% 
  inner_join( distinct(obs_filtered_age, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
  distinct(patient_num, n_hospitalisation) %>% 
  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
# Adding new columns
observation_table <- obs_filtered_age %>% 
  inner_join(clinical_table[, c("patient_num", "days_since_admission")], by = c("patient_num", "days_since_admission"))
demographic_table <- demo_raw %>% 
  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
all_codes <- observation_table %>% 
   inner_join(select(clinical_table, -calendar_date), by = c("patient_num", "days_since_admission")) %>% 
   inner_join(demographic_table[, c("patient_num", "sex")], by = "patient_num")
```

**Create Lurie's Input_data table**

```{r}
nrow(unique(data.frame(clinical_filtered$patient_num)))
nrow(clinical_filtered %>% distinct(patient_num, encounter_num))

input_data <- clinical_filtered %>% 
  mutate(age_years = floor(age_years)) %>% 
  filter(date >= start_date_plots & date <= end_date_plots,
         age_years > 10 & age_years < 18) %>% 
  group_by(patient_num, encounter_num) %>% 
  arrange(date) %>% 
  slice(1L) %>% 
  ungroup() %>% 
  select(patient_num, encounter_num, race, ethnicity, gender, age_years, date, i_hsp_dsch_dttm, weeks, month, year, period, time_p, n_hospitalisation, len_hospitalisation) %>% 
  rename("sex" = gender,
         "age" = age_years)

nrow(unique(data.frame(input_data$patient_num)))
nrow(input_data %>% distinct(patient_num, encounter_num))
```

```{r eval=FALSE, include=FALSE}
qc_check <- admit_all_codes %>% 
   rename("patient_num" = patient_id, 
          "concept_code" = diagnosis_code) %>% 
  left_join(., icdCodes, by = c("concept_code" = "ICD10_Code")) %>% 
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others")) %>% 
  left_join(input_data, ., by = "patient_num") %>% 
  mutate(i_diagnosis_dttm = strptime(diagnosis_dttm, format="%m/%d/%Y") %>% as.POSIXct(),
         date = format(as.POSIXct(date, format = "%m/%d/%Y %H:%M"), format = "%Y-%m-%d"),
         i_hsp_dsch_dttm = format(as.POSIXct(i_hsp_dsch_dttm, format = "%m/%d/%Y %H:%M"), format = "%Y-%m-%d")) %>% 
  dplyr::filter( diagnosis_code_source == "ICD-10-CM", 
                 #in_hospital == 1,
                 i_diagnosis_dttm >= date & i_diagnosis_dttm <= i_hsp_dsch_dttm,
                 age > 10 & age < 18,
                 date >= start_date_plots & date <= end_date_plots,
                 ) %>%
  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category)

```{r eval=FALSE, include=FALSE}
stopifnot(unique(all_codes$cohort) == "AllAdm")
stopifnot(unique(all_codes$in_hospital) == 1)
stopifnot(unique(all_codes$len_hospitalisation >= min_hosp_days))
stopifnot(nrow(all_codes) == nrow(observation_table))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(demographic_table$patient_num)))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(observation_table$patient_num)))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(clinical_table$patient_num)))
stopifnot(nrow(distinct(all_codes, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table, patient_num, n_hospitalisation)))
stopifnot(nrow(distinct(all_codes, patient_num, days_since_admission)) == nrow(distinct(observation_table, patient_num, days_since_admission)))
```

@@

```{r eval=FALSE, include=FALSE}
# Filtering hospitalizations where patient age is in inclusion criteria
observation_table_ado <- filter(observation_table, age_time_diagnosis >= 11 & age_time_diagnosis <= 17)

@@

```{r eval=FALSE, include=FALSE}
stopifnot(unique(all_codes_ado$cohort) == "AllAdm")
stopifnot(unique(all_codes_ado$in_hospital) == 1)
stopifnot(unique(all_codes_ado$len_hospitalisation >= min_hosp_days))
stopifnot(nrow(all_codes_ado) == nrow(observation_table_ado))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(demographic_table_ado$patient_num)))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(observation_table_ado$patient_num)))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(clinical_table_ado$patient_num)))
stopifnot(nrow(distinct(all_codes_ado, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table_ado, patient_num, n_hospitalisation)))
stopifnot(nrow(distinct(all_codes_ado, patient_num, days_since_admission)) == nrow(distinct(observation_table_ado, patient_num, days_since_admission)))
```

Are there any patients in the admit_demog df who do not have any icd codes listed?

Note: this looks fine
```{r}
admit_demog <- read.csv(paste0(folder_4ce_files, "/psych_admit_demog.csv"))

# list of all patients who should have had at least 1 psych diagnosis
psych_ids <- admit_demog %>% 
  distinct(patient_id)
nrow(psych_ids)

# evaluate whether there are patients in admit_all_codes file that were not listed in the admit_demog file
admit_demog_ids <- admit_all_codes %>% 
  distinct(patient_id) %>% 
  filter(!patient_id %in% psych_ids$patient_id)
nrow(admit_demog_ids)

# evaluate whether there are patients in the admit_demog file that do not have diagnosis codes
psych_id_check <- psych_ids %>% 
  distinct(patient_id) %>% 
  filter(!patient_id %in% admit_all_codes$patient_id)
nrow(psych_id_check)
```

```{r message=FALSE, warning=FALSE}
# Getting site name, will be used later in the analysis
#site <- unique(as.character(demo_raw$siteid))
site <- "Lurie"

# all_codes_desc <- all_codes_ado %>% 
#     left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
#   mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
#                                "psy", 
#                                "others")) %>%
#   dplyr::select( patient_num, sex, age_time_diagnosis, concept_type, concept_code, calendar_date,
#                  in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
#                  period, time_p, disorder_group, description, icd_code_category )

# begin processing the ICD codes for the patients
all_codes_desc <- admit_all_codes %>% 
   rename("patient_num" = patient_id, 
          "concept_code" = icd_10_code) %>% 
  left_join(., icdCodes, by = c("concept_code" = "ICD10_Code")) %>% 
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others"))

nrow(input_data)
nrow(unique(data.frame(input_data$patient_num)))
nrow(input_data %>% distinct(patient_num, encounter_num))

# Join mental health codes to the patient_number
input_data <- input_data %>% 
  left_join(., all_codes_desc, by = "patient_num") %>% 
  mutate(i_diagnosis_dttm = strptime(diagnosis_dttm, format="%m/%d/%Y") %>% as.POSIXct(),
         date = strptime(date, format="%Y-%m-%d") %>% as.POSIXct(),
         i_hsp_dsch_dttm = strptime(i_hsp_dsch_dttm, format="%Y-%m-%d") %>% as.POSIXct()) %>% 
  dplyr::filter(date >= start_date_plots & date <= end_date_plots)

nrow(input_data)
nrow(unique(data.frame(input_data$patient_num)))
nrow(input_data %>% distinct(patient_num, encounter_num))

# we will temporarily split the data frame between our psych patients with ICD codes and our non-pysch without, due to the filtering function automatically dropping rows with missing ICD codes
# of note: these are just some patients who have an ICD code - could be some non-psych codes as the original query including some icd codes for poisoning that ended up being excluded
keep_psych <- input_data %>% 
  group_by(patient_num, encounter_num) %>% 
  filter(i_diagnosis_dttm >= date & i_diagnosis_dttm <= i_hsp_dsch_dttm) %>% 
  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category)

keep_non_psych = input_data %>% 
  group_by(patient_num, encounter_num) %>% 
  filter(is.na(i_diagnosis_dttm)) %>% 
  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category)

# keep non_psych whose dx date did not occur during the admission
# but correct the diagnosis data to match the non_psych group
keep_non_psych_v2 = input_data %>% 
  group_by(patient_num, encounter_num) %>% 
  filter(i_diagnosis_dttm < date | i_diagnosis_dttm >= i_hsp_dsch_dttm) %>% 
  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category) %>% 
  mutate(concept_code = NA,
         disorder_group = NA,
         description = NA, 
         icd_code_category = NA)

input_data <- rbind(keep_psych, keep_non_psych, keep_non_psych_v2)

nrow(input_data)
nrow(unique(data.frame(input_data$patient_num)))
nrow(input_data %>% distinct(patient_num, encounter_num))

all_codes_desc <- input_data %>% 
  mutate( icd_code_letter = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))

all_codes_desc <- left_join( all_codes_desc, icdCategory) %>%
  mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others"),
          icd_code_category = ifelse(is.na(icd_code_category), "others", icd_code_category))

nrow(all_codes_desc %>% distinct(patient_num, encounter_num))

all_codes_psy <- all_codes_desc %>% 
  filter(concept_code %in% icdCodes$ICD10_Code)

nrow(unique(data.frame(all_codes_psy$patient_num)))
nrow(all_codes_psy %>% distinct(patient_num, encounter_num))
# View(all_codes_psy %>% 
#        ungroup() %>% 
#        distinct(patient_num, encounter_num, time_p, period) %>% 
#        group_by(period) %>% 
#        mutate(count = n_distinct(patient_num, encounter_num)) %>% 
#        distinct(period, count))

## how many weeks in 'during' vs 'before' pandemic (15 for both)
# View(all_codes_psy %>% 
#        ungroup() %>% 
#        distinct(month, period) %>%
#        group_by(period) %>% 
#        mutate(count = n_distinct(month)))

#all_codes_desc$i_date <- strptime(all_codes_desc$date, format="%Y-%m-%d") %>% as.POSIXct()

# if( raceAvailable == TRUE ){
#   all_codes_psy <- all_codes_psy %>%
#     left_join( race_raw, by = c("patient_num") )
# }
```

Evaluate numbers before vs after pandemic

```{r}
before_v_after <- all_codes_desc %>% 
  distinct(patient_num, encounter_num, period, icd_code_category) %>% 
  group_by(period, icd_code_category) %>% 
  mutate(count = n_distinct(patient_num, encounter_num)) %>% 
  distinct(period, icd_code_category, count)

before_v_after
```

QC Checks for Psych patients

```{r}
p1 <- ggplot(all_codes_desc %>%
               filter(icd_code_category == "psy") %>% 
               group_by(weeks, period) %>% 
               arrange(weeks) %>% 
               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
                      weeks = as.Date(weeks)) %>% 
               distinct(weeks, psych_any_count, period),
       aes(x = weeks, y = psych_any_count, fill = period)) + 
  geom_bar(stat="identity", binwidth=1) + 
  geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
  scale_color_brewer(palette="Dark2", direction = -1) +
  scale_fill_brewer(palette = "Dark2", direction = -1) + 
  ylab("Any Psychiatric Disease") + 
  xlab("") + 
  #scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
  theme_bw() + 
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))

p1
